---
title: "`r params$title`"
author: "`r params$author`"
date: "`r format(Sys.Date(), '%d %B %Y')`"
params:
  results: NULL
  title: "Stock assessment - stockflow"
  author: "stockflow"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
    df_print: kable
  pdf_document:
    toc: true
    number_sections: true
  word_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 8,
  fig.height = 4.8,
  dpi = 150
)

res <- params$results

# Availability of sections
has_growth    <- !is.null(res$growth)
has_mortality <- !is.null(res$mortality)
has_allometry <- !is.null(res$allometry)
has_maturity  <- !is.null(res$maturity)
has_lbb       <- !is.null(res$lbb)
has_lbspr     <- !is.null(res$lbspr)
has_ypr       <- !is.null(res$ypr)
has_tac       <- !is.null(res$tac)
has_quota     <- !is.null(res$quota)
has_closure   <- !is.null(res$closure)
has_size      <- !is.null(res$size_advice)
has_mse       <- !is.null(res$mse)
has_gg        <- requireNamespace("ggplot2", quietly = TRUE)

kb <- function(x, ...) knitr::kable(x, row.names = FALSE, digits = 3, ...)
```

# Executive Summary

```{r meta}
meta <- res$meta
if (length(meta)) {
  kb(data.frame(Champ = names(meta),
                Value = vapply(meta, function(v) paste(v, collapse = ", "),
                                character(1)),
                stringsAsFactors = FALSE))
}
```

```{r summary-table}
st <- tryCatch(
  if (requireNamespace("stockflow", quietly = TRUE))
    stockflow::summary_table(res)
  else
    summary_table(res),
  error = function(e) NULL
)
if (!is.null(st) && nrow(st) > 0) kb(st) else cat("No indicators available.\n")
```

```{r conclusion, results='asis'}
# Summary diagnostic
statuts <- c()
if (has_lbspr && nrow(res$lbspr$summary) > 0)
  statuts <- c(statuts,
               paste("LBSPR:", res$lbspr$summary$statut[nrow(res$lbspr$summary)]))
if (has_lbb && !is.null(res$lbb$ref_levels))
  statuts <- c(statuts, paste("LBB:", res$lbb$ref_levels$statut))
if (has_mortality)
  statuts <- c(statuts, paste("Exploitation:", res$mortality$summary$statut))

if (length(statuts))
  cat("\n**Diagnostic:**\n\n", paste0("- ", statuts, collapse = "\n"), "\n")
```

```{r growth-title, eval=has_growth, results='asis'}
cat("\n# Growth\n")
```

```{r growth-tab, eval=has_growth}
g <- res$growth
if (!is.null(g$comparison)) {
  cat("Multi-criteria comparison of fitted methods:\n")
  kb(g$comparison)
}
```

```{r growth-best, eval=has_growth, results='asis'}
if (!is.null(g$best))
  cat(sprintf("\n**Selected model: %s** — Linf = %.2f, K = %.3f, phiL = %.3f\n",
              g$best$method, g$best$Linf, g$best$K, g$best$phiL))
```

```{r growth-curve, eval=has_growth && has_gg}
b <- g$best
if (!is.null(b) && is.finite(b$Linf) && is.finite(b$K)) {
  ages <- seq(0, 6, by = 0.05)
  L    <- b$Linf * (1 - exp(-b$K * (ages - 0)))
  df   <- data.frame(age = ages, length = L)
  ggplot2::ggplot(df, ggplot2::aes(age, length)) +
    ggplot2::geom_line(colour = "#12507b", linewidth = 1) +
    ggplot2::labs(title = "Von Bertalanffy curve",
                  x = "Age (year)", y = "Length") +
    ggplot2::theme_bw()
}
```

```{r mortalite-titre, eval=has_mortality, results='asis'}
cat("\n# Mortality\n")
```

```{r mortalite-synth, eval=has_mortality, results='asis'}
mo <- res$mortality
cat(sprintf("\n- Z = %.3f\n- M (consensus) = %.3f\n- F = %.3f\n- E = F/Z = %.3f\n- F/M = %.3f\n- **Status: %s**\n",
            mo$Z$Z, mo$M_consensus, mo$summary$F,
            mo$summary$E, mo$summary$FM, mo$summary$statut))
```

```{r mortalite-M, eval=has_mortality}
cat("\nEstimates of M by method (TropFishR, FSA, fishmethods, Jensen):\n")
kb(res$mortality$comparison)
```

```{r mortalite-plot, eval=has_mortality && has_gg}
cmp <- res$mortality$comparison
if (nrow(cmp) > 1) {
  cmp$method <- factor(cmp$method, levels = cmp$method[order(cmp$M)])
  ggplot2::ggplot(cmp, ggplot2::aes(M, method)) +
    ggplot2::geom_point(colour = "#12507b", size = 2) +
    ggplot2::geom_vline(xintercept = res$mortality$M_consensus,
                        linetype = 2, colour = "#c0392b") +
    ggplot2::labs(title = "Natural mortality M by method",
                  subtitle = "Red line = consensus (geometric mean)",
                  x = "M", y = NULL) +
    ggplot2::theme_bw()
}
```

```{r allometrie-titre, eval=has_allometry, results='asis'}
cat("\n# Length-weight allometry\n")
```

```{r allometrie, eval=has_allometry, results='asis'}
a <- res$allometry
cat(sprintf("\nRelationship: **W = %.5g · L^%.3f** (r² = %.3f, n = %d)\n\n95%% CI for b: [%.3f ; %.3f] — %s\n",
            a$a, a$b, a$r2, a$n, a$b_ci[1], a$b_ci[2], a$allometry))
```

```{r maturite-titre, eval=has_maturity, results='asis'}
cat("\n# Maturity\n")
m <- res$maturity
cat(sprintf("\n- L50 = %.2f\n- L95 = %.2f\n", m$L50, m$L95))
```

```{r lbb-titre, eval=has_lbb, results='asis'}
cat("\n# LBB — Length-Based Bayesian Biomass\n")
```

```{r lbb-tab, eval=has_lbb}
kb(res$lbb$summary)
```

```{r lbb-plot, eval=has_lbb && has_gg}
df <- res$lbb$summary
long <- data.frame(
  Year = rep(df$Year, 3),
  ind  = rep(c("B/B0", "B/Bmsy", "F/M"), each = nrow(df)),
  val  = c(df$BB0, df$BBmsy, df$FM)
)
long <- long[is.finite(long$val), ]
if (nrow(long) > 0) {
  ggplot2::ggplot(long, ggplot2::aes(Year, val)) +
    ggplot2::geom_line(colour = "#12507b") +
    ggplot2::geom_point(size = 1) +
    ggplot2::geom_hline(yintercept = 1, linetype = 2, colour = "grey40") +
    ggplot2::facet_wrap(~ ind, scales = "free_y") +
    ggplot2::labs(title = "LBB — status indicators", x = NULL, y = NULL) +
    ggplot2::theme_bw()
}
```

```{r lbspr-titre, eval=has_lbspr, results='asis'}
cat("\n# LBSPR — Spawning Potential Ratio (SPR)\n")
```

```{r lbspr-tab, eval=has_lbspr}
kb(res$lbspr$summary)
```

```{r lbspr-plot, eval=has_lbspr && has_gg}
df  <- res$lbspr$summary
tgt <- res$lbspr$targets
ggplot2::ggplot(df, ggplot2::aes(Year, SPR)) +
  ggplot2::geom_ribbon(ggplot2::aes(ymin = 0, ymax = tgt[["SPR_limit"]]),
                       fill = "#FF7F7F", alpha = .2) +
  ggplot2::geom_line(colour = "#12507b") +
  ggplot2::geom_point(size = 1.5) +
  ggplot2::geom_hline(yintercept = tgt[["SPR_target"]], linetype = 2,
                      colour = "darkgreen") +
  ggplot2::geom_hline(yintercept = tgt[["SPR_limit"]], linetype = 3,
                      colour = "red") +
  ggplot2::ylim(0, 1) +
  ggplot2::labs(title = "SPR by year",
                subtitle = "Green = target; red = limit",
                x = NULL, y = "SPR") +
  ggplot2::theme_bw()
```

```{r ypr-titre, eval=has_ypr, results='asis'}
cat("\n# YPR — Yield per recruit and reference points\n")
```

```{r ypr-tab, eval=has_ypr}
kb(res$ypr$reference_points)
```

```{r ypr-txt, eval=has_ypr, results='asis'}
rp <- res$ypr$reference_points[1, , drop = FALSE]
if (!is.null(rp$F01) && !is.null(rp$Fmax))
  cat(sprintf(paste0("\nPer-recruit reference points (Beverton & Holt): ",
                     "**F0.1 = %.3f**, Fmax = %.3f, E0.1 = %.3f, Emax = %.3f.\n",
                     "\nF0.1 is the recommended precautionary management point ",
                     "(slope = 10%% of the slope at the origin).\n"),
              rp$F01, rp$Fmax, rp$E01, rp$Emax))
```

```{r tac-titre, eval=has_tac, results='asis'}
cat("\n# TAC — Total Allowable Catch\n")
```

```{r tac-txt, eval=has_tac, results='asis'}
tc <- res$tac
cat(sprintf(paste0("\n- Method: %s\n- Raw TAC: %.1f t\n",
                   if (tc$buffer > 0) "- Precaution: -%.0f%%\n" else "%s",
                   "- **Final TAC: %.1f t**\n"),
            tc$method, tc$TAC_raw,
            if (tc$buffer > 0) 100 * tc$buffer else "",
            tc$TAC))
```

```{r quota-tab, eval=has_quota, results='asis'}
cat("\n## Quota allocation by segment / fleet\n\n")
kb(res$quota$quotas)
cat(sprintf("\nSum of quotas: %.1f t (difference of %.2f t from the TAC).\n",
            res$quota$somme_quotas, res$quota$ecart_somme))
```

```{r quota-plot, eval=has_quota && has_gg}
q <- res$quota$quotas
q$segment <- factor(q$segment, levels = q$segment[order(q$quota_t)])
ggplot2::ggplot(q, ggplot2::aes(quota_t, segment)) +
  ggplot2::geom_col(fill = "#12507b") +
  ggplot2::geom_text(ggplot2::aes(label = sprintf("%.0f t", quota_t)),
                     hjust = -0.1, size = 3) +
  ggplot2::labs(title = "Quota allocation by segment",
                x = "Quota (t)", y = NULL) +
  ggplot2::expand_limits(x = max(q$quota_t) * 1.12) +
  ggplot2::theme_bw()
```

```{r closure-titre, eval=has_closure, results='asis'}
cat("\n# Biological rest period — optimal window and duration\n")
```

```{r closure-txt, eval=has_closure, results='asis'}
b <- res$closure$best
cat(sprintf(paste0("\n**Recommended window: %d months (%s).** ",
                   "It protects %.0f%% of the annual reproduction and recruitment score.\n"),
            b$duration, b$months_label, 100 * b$protected))
```

```{r closure-tab, eval=has_closure}
if (!is.null(res$closure$by_duration)) {
  cat("Comparison of closure durations:\n")
  kb(res$closure$by_duration)
}
```

```{r closure-plot, eval=has_closure && has_gg}
df <- res$closure$monthly
prot <- res$closure$best$months
df$statut <- ifelse(df$month %in% prot, "Proposed window", "Outside closure")
df$month_label <- factor(df$month_label, levels = df$month_label)
ggplot2::ggplot(df, ggplot2::aes(month_label, score, fill = statut)) +
  ggplot2::geom_col() +
  ggplot2::scale_fill_manual(values = c("Proposed window" = "#c0392b",
                                        "Outside closure" = "grey75"), name = NULL) +
  ggplot2::labs(title = "Monthly closure score and optimal window",
                x = NULL, y = "Closure score") +
  ggplot2::theme_bw() +
  ggplot2::theme(legend.position = "bottom")
```

```{r size-titre, eval=has_size, results='asis'}
cat("\n# Technical measures — management sizes\n")
```

```{r size-tab, eval=has_size, results='asis'}
r <- res$size_advice$reference
cat(sprintf("\nReference sizes: Lopt = %.1f cm, Lc_opt = %.1f cm%s.\n\n",
            r$Lopt, r$Lc_opt,
            if (!is.null(r$L50_maturity))
              sprintf(", L50 maturity = %.1f cm", r$L50_maturity) else ""))
kb(res$size_advice$advice)
```

```{r mse-titre, eval=has_mse, results='asis'}
cat("\n# Management measures — scenario comparison\n")
```

```{r mse-tab, eval=has_mse}
kb(res$mse$comparison)
```

```{r mse-plot, eval=has_mse && has_gg}
df  <- res$mse$comparison
tgt <- res$mse$targets
ggplot2::ggplot(df, ggplot2::aes(SPR, Yield_rel, label = scenario)) +
  ggplot2::annotate("rect", xmin = 0, xmax = tgt[["SPR_limit"]],
                    ymin = -Inf, ymax = Inf, fill = "#FF7F7F", alpha = .18) +
  ggplot2::geom_vline(xintercept = tgt[["SPR_target"]], linetype = 2,
                      colour = "darkgreen") +
  ggplot2::geom_point(size = 3, colour = "#12507b") +
  ggplot2::geom_text(vjust = -0.8, size = 3) +
  ggplot2::labs(title = "Trade-off between yield and spawning potential",
                x = "SPR", y = "Relative yield") +
  ggplot2::theme_bw()
```

```{r mse-reco, eval=has_mse, results='asis'}
df <- res$mse$comparison
ok <- df[is.finite(df$SPR) & df$SPR >= res$mse$targets[["SPR_target"]], ]
if (nrow(ok) > 0) {
  best <- ok[which.max(ok$Yield_rel), ]
  cat(sprintf("\n**Recommendation:** among the scenarios reaching the SPR target, *%s* maximizes relative yield (SPR = %.2f, yield = %.3f).\n",
              best$scenario, best$SPR, best$Yield_rel))
} else {
  cat("\n**Alert:** none of the tested scenarios reach the SPR target — strengthened measures are needed.\n")
}
```

# Methods and References

Analyses carried out with the **stockflow** package, which wraps the
official functions of:

- **TropFishR** — Mildenberger, T.K., Taylor, M.H., Wolff, M. (2017). *Methods Ecol. Evol.* 8(11).
- **FSA** — Ogle, D.H. (2016). *Introductory Fisheries Analyses with R*.
- **LBB** — Froese, R., Winker, H., Coro, G. et al. (2018). *ICES J. Mar. Sci.* 75(6): 2004-2015.
- **LBSPR** — Hordyk, A. et al. (2015). *ICES J. Mar. Sci.* 72(1): 217-231.
- **openMSE / MSEtool** — Carruthers, T.R., Hordyk, A.R. (2018). *Methods Ecol. Evol.* 9(12).

Natural mortality estimators: Pauly (1980), Hoenig (1983), Jensen (1996),
Gislason et al. (2010), Then et al. (2015).

```{r session, results='asis'}
cat(sprintf("\n---\n\n*Report generated on %s with stockflow.*\n",
            format(Sys.time(), "%d/%m/%Y at %H:%M")))
```