---
title: "Stockflow Tutorial — from raw file to stock assessment"
author: "Kamarel Ba"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
vignette: >
  %\VignetteIndexEntry{Tutoriel stockflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 5,
  collapse = TRUE,
  comment = "#>"
)

# Detection of package availability (heavy chunks are conditional)
has_tropfish <- requireNamespace("TropFishR", quietly = TRUE)
has_fsa      <- requireNamespace("FSA",       quietly = TRUE)
has_jags     <- requireNamespace("R2jags",    quietly = TRUE)
has_lbspr    <- requireNamespace("LBSPR",     quietly = TRUE)
```

# Purpose of the tutorial

This tutorial walks through **the entire `stockflow` pipeline** on the example (fictitious) data
bundled with the package. The running example is
the volute *Cymbium cymbium*; the same functions apply to octopus
(*Octopus vulgaris*) and shrimp (*Penaeus notialis*), also provided.

The complete pipeline:

```
Import / Data ---> Graphical exploration ---> LFQ preparation
        |                                              |
   Allometry <---------- maturity (L50) ----------> Growth (VBGF)
        |                     |                        |
        +--------> Mortality (M, Z, F, E) <-----------+
                              |
        +---------------------+---------------------+
        |                     |                     |
   LBB / LBSPR            YPR / SPR            Biological rest period
        |                     |                     |
        +--> TAC + quotas <---+     Technical measures (Lc, mesh size)
                              |
                     Automatic report HTML / PDF / DOCX
```

The package follows a strict rule: **it never implements a scientific
algorithm itself**. Each function is a documented *wrapper* around the
official implementations of **TropFishR**, **FSA**, **LBSPR**, **LBB**, and
**fishmethods**.

> **Reproducibility.** No function uses `library()`; everything goes through
> `stockflow::function()` or `package::function()`. The costly steps (ELEFAN,
> LBB via JAGS) are shown but not executed when the vignette is built
> (`eval = FALSE`).

# Setup

```{r load-package}
library(stockflow)
```

The datasets are available as soon as the package is loaded (`LazyData`):

```{r data-overview}
data("cymbium_bio",   package = "stockflow")
data("cymbium_freq",  package = "stockflow")

# Working on a single species for the running example
bio  <- subset(cymbium_bio,  species == "Cymbium cymbium")
freq <- subset(cymbium_freq, species == "Cymbium cymbium")

c(n_biologie = nrow(bio), n_frequences = nrow(freq),
  LCQ_min = min(freq$LCQ, na.rm = TRUE),
  LCQ_max = max(freq$LCQ, na.rm = TRUE))
```

Datasets provided with the package:

| Dataset            | Description                              | Rows    |
|--------------------|------------------------------------------|---------|
| `cymbium_bio`      | Volute biological data (4 species)       | 3 130   |
| `cymbium_freq`     | Volute size frequencies                  | 8 243   |
| `octopus_bio`      | Octopus biological data                  | 1 682   |
| `octopus_freq`     | Octopus size frequencies                 | 1 689   |
| `penaeus_bio`      | Shrimp biological data (+ gonads)        | 6 286   |
| `penaeus_freq`     | Shrimp size frequencies                  | 31 043  |
| `total_catches` | Annual catches by group                     | 55      |
| `annual_cpue`    | Annual CPUE by group                        | 47      |

# Step 1 — Graphical exploration

The graphical module (`plot_*`) provides publication-ready figures, all
paired with **confidence intervals** where relevant.

## Size distribution

```{r plot-lf}
plot_length_frequency(freq, length_col = "LCQ", sex_col = "sex")
```

## Length-weight relationship (power AND log-log)

```{r plot-lw, fig.width=10}
plot_length_weight(bio, length_col = "LCQ", weight_col = "weight",
                   scale = "both")
```

## Monthly size evolution

Useful for tracking the modal progression of cohorts:

```{r plot-evolution}
plot_length_evolution(freq, length_col = "LCQ", date_col = "Date",
                      period = "month", type = "density")
```

```{r plot-evolution-hist}
plot_length_evolution(freq, length_col = "LCQ", date_col = "Date",
                      period = "month", type = "histogram")
```

# Step 2 — Preparing the LFQ object

`prepare_tropfish()` builds an `lfq` object compatible with **TropFishR** from
`Length`, `Year`, `Month` columns.

```{r prepare-lfq, eval=has_tropfish}
freq$Length <- freq$LCQ
freq$Year   <- as.integer(format(freq$Date, "%Y"))
freq$Month  <- as.integer(format(freq$Date, "%m"))

lfq <- prepare_tropfish(freq, bin_size = 2, species = "Cymbium cymbium",
                        stock_name = "Cymbium", length_unit = "cm")
c(classe = class(lfq), n_classes = length(lfq$midLengths),
  n_dates = length(lfq$dates))
```

# Step 3 — Growth (von Bertalanffy)

`stockflow` follows the **TropFishR growth protocol** in four stages:

1. **Powell-Wetherall** — a first estimate of \(L_{inf}\) (and of \(Z/K\));
2. **K-Scan** — with \(L_{inf}\) fixed, \(K\) is scanned to find the value that
   maximizes the restructuring score \(R_n\) (informed starting value);
3. **ELEFAN** — search over the **response surface** \(L_{inf}\times K\)
   (grid), then refinement via **simulated annealing** (SA) and **genetic
   algorithm** (GA);
4. **Multi-criteria comparison** and automatic selection of the best model,
   with visual verification: the growth curve must **pass through the
   modes** of the restructured frequencies.

## Powell-Wetherall then K-Scan

`run_powell()` calls `TropFishR::powell_wetherall()`. The regression interval
`reg_int` is normally requested interactively; in a script it is provided
explicitly.

```{r growth-powell, eval=has_tropfish}
pw <- run_powell(lfq, reg_int = c(2, 12))
ks <- run_kscan(lfq, Linf = pw, K_range = seq(0.1, 1.5, 0.1))
c(Linf_Powell = round(ks$Linf, 2), K_optimal = round(ks$K, 3))
```

```{r kscan-plot, eval=has_tropfish}
plot_response_surface(ks)   # Rn(K) score profile at fixed Linf
```

## ELEFAN response surface

`run_elefan()` wraps `TropFishR::ELEFAN()`: the grid search produces a score
matrix plotted as a heatmap with a confidence contour.

```{r elefan-grid, eval=has_tropfish}
el <- run_elefan(lfq, Linf_range = seq(36, 46, 2),
                 K_range = seq(0.2, 1.0, 0.2), MA = 5)
plot_response_surface(el)
```

Refinement via simulated annealing and genetic algorithm is more costly (1–3
min); it is shown without being executed, along with the multi-criteria
comparison and automatic selection.

```{r growth-elefan-refine, eval=FALSE}
ga  <- run_elefan_ga(lfq, Linf_range = c(36, 46), K_range = c(0.2, 1.0))
sa  <- run_elefan_sa(lfq, Linf_range = c(36, 46), K_range = c(0.2, 1.0))

# Orchestrator: Powell + K-Scan + ELEFAN(grid/GA/SA) + comparison + selection
an <- run_growth_analysis(lfq, Linf_range = c(36, 46), K_range = c(0.2, 1.0))
an$comparison   # multi-criteria table (Rn_max, phi', ...)
an$best         # best model selected
```

## Verification: does the curve pass through the modes?

`plot_growth_modes()` overlays the VBGF curve on the **restructured**
frequencies (positive scores = modes in blue). This is the final visual check
of the protocol.

```{r growth-modes, eval=has_tropfish, fig.height=5}
plot_growth_modes(lfq, el, MA = 5)
```

For the rest of the tutorial, plausible parameters are fixed for *Cymbium
cymbium*:

```{r growth-fixed}
Linf <- 42;  K <- 0.45;  t0 <- 0
```

## Growth curve (age → length)

```{r vbgf-curve, eval=has_tropfish}
plot_vbgf(Linf = Linf, K = K, t0 = t0, tmax = 8, length_unit = "cm")
```

# Step 4 — Mortality

## Natural mortality M — all methods

`estimate_M_all()` combines the empirical estimators from **FSA**
(`Mmethods` + `metaM`, ~25 methods), from **TropFishR**, and from Jensen, then
derives a **consensus** (geometric mean).

```{r mortality-M}
M <- estimate_M_all(Linf = Linf, K = K, temp = 24, tmax = 8)
c(n_methodes = nrow(M$table),
  M_geomean  = round(M$consensus[["geomean"]], 3),
  M_median   = round(M$consensus[["median"]], 3))
```

```{r mortality-M-plot}
plot(M)
```

## Total mortality Z, then F and E

In production, `run_mortality()` chains the length-converted catch curve (Z),
the M consensus, then F = Z − M and the exploitation rate E = F / Z. The catch
curve requires a regression interval (interactive if not supplied); it is
therefore shown without being executed.

```{r mortality-run, eval=FALSE}
mort <- run_mortality(lfq, growth_model = list(Linf = Linf, K = K, t0 = t0),
                      temp = 24, tmax = 8)
mort$summary   # columns Z, M, F, E
```

# Step 5 — Length-weight allometry

```{r allometry}
lw <- fit_length_weight(bio, length_col = "LCQ", weight_col = "weight")
c(a = round(lw$a, 4), b = round(lw$b, 3),
  b_ic_bas = round(lw$b_ci[1], 3), b_ic_haut = round(lw$b_ci[2], 3),
  R2 = round(lw$r2, 3))
```

The confidence interval on *b* allows testing for isometry (b = 3):

```{r allometry-plot}
plot(lw)
```

# Step 6 — Maturity (L50)

The logistic maturity ogive gives the size at first maturity L50, essential
for LBSPR and for technical measures. The `plot_maturity_ogive()` function
plots the ogive with a CI (delta method on L50).

```{r maturity-plot}
plot_maturity_ogive(bio, length_col = "LCQ", maturity_col = "maturity",
                    mature_codes = c("M", "P"))
```

L50 / L95 are recovered numerically via the binomial GLM:

```{r maturity-values}
d <- bio[!is.na(bio$maturity) & !is.na(bio$LCQ), ]
d$mature <- as.integer(d$maturity %in% c("M", "P"))
gm  <- stats::glm(mature ~ LCQ, data = d, family = stats::binomial())
cf  <- stats::coef(gm)
L50 <- as.numeric(-cf[1] / cf[2])
L95 <- as.numeric((log(0.95 / 0.05) - cf[1]) / cf[2])
c(L50 = round(L50, 2), L95 = round(L95, 2))
```

# Step 7 — LBSPR (Spawning Potential Ratio)

```{r lbspr, eval=has_lbspr}
pars <- lbspr_pars(Linf = Linf, L50 = L50, L95 = L95, MK = 1.5)
# selectivity and F/M on the LB_pars object
pars@SL50 <- L50;  pars@SL95 <- L95;  pars@FM <- 1
sim <- lbspr_sim(pars)
c(SPR = round(sim@SPR, 3))
```

# Step 8 — LBB (Length-Based Bayesian Biomass)

The LBB method (Froese's official script, via **JAGS**) estimates B/B0, Y/R,
and reference points. It requires the external JAGS software; the workflow is
shown without being executed.

```{r lbb, eval=FALSE}
lbb_in  <- prepare_lbb_data(lfq)
lbb_out <- run_lbb_froese(lbb_in, stock = "Cymbium")
extract_lbb(lbb_out)
```

The length-based reference points derived from Linf are directly accessible:

```{r lbb-refpts}
rp_lb <- lbb_reference_points(Linf = Linf, MK = 1.5, FM = 1)
c(Lopt = round(rp_lb$Lopt, 1), Lc_opt = round(rp_lb$Lc_opt, 1))
```

# Step 9 — YPR and per-recruit reference points

`ypr_param()` assembles the parameters, `run_ypr()` runs TropFishR's
per-recruit model, `ypr_reference_points()` extracts F0.1, Fmax, and the
associated E values.

```{r ypr}
Mstar <- M$consensus[["geomean"]]
yp <- ypr_param(Linf = Linf, K = K, t0 = t0, M = Mstar,
                a = lw$a, b = lw$b, Lc = L50, Lr = 4, bin_size = 2)
yr <- run_ypr(yp)
rp <- ypr_reference_points(yr)
c(F0.1 = round(rp$F01, 3), Fmax = round(rp$Fmax, 3),
  E0.1 = round(rp$E01, 3), Emax = round(rp$Emax, 3))
```

```{r ypr-plot}
plot(yr)
```

# Step 10 — TAC and quota allocation

The TAC is derived from the current catch and an exploitation target (here
E0.1), with a precautionary buffer. Quotas are then allocated by species in
proportion to historical catches.

```{r tac}
tac <- compute_tac(method = "ratio_E", catch_current = 1200,
                   E_current = 0.73, E_target = rp$E01, buffer = 0.10)
c(TAC_brut = round(tac$TAC_raw, 1), TAC_final = round(tac$TAC, 1))
```

```{r quota}
# Pro-rata of the 4 Cymbium species (proxy = sampled biomass)
agg <- aggregate(weight ~ species, data = cymbium_bio,
                 FUN = function(x) sum(x, na.rm = TRUE) / 1000)
names(agg) <- c("espece", "capture_t")
agg$annee  <- 2023

pr <- compute_prorata(agg, segment_col = "espece",
                      catch_col = "capture_t", year_col = "annee")
q  <- allocate_quota(tac, pr)
q$quotas
```

# Step 11 — Biological rest period (spatio-temporal closure)

`spawning_season()` computes the monthly proportion of spawners;
`optimal_closure()` searches for the closure window that best protects
reproduction (circular sliding window).

```{r closure}
sp  <- spawning_season(bio, date_col = "Date", maturity_col = "maturity",
                       mature_codes = c("M", "P"))
clo <- optimal_closure(spawning = sp, duration = 3)
c(fenetre = clo$best$months_label,
  duree_mois = length(clo$best$months),
  part_protegee = round(clo$best$protected, 3))
```

```{r closure-plot}
plot(clo)
```

# Step 12 — Technical measures (management size, mesh size)

`recommend_sizes()` evaluates candidate minimum sizes against L50, Lopt, and
Lc_opt; `retention_curve()` plots the gear's retention ogive.

```{r sizes}
adv <- recommend_sizes(candidates = c(8, 10, 12, 14),
                       Linf = Linf, L50_maturity = L50, MK = 1.5)
adv$advice[, c("Lc", "verdict")]
```

```{r retention}
ret <- retention_curve(Lt = seq(2, 20, 0.5), type = "trawl_ogive",
                       L50 = L50, L75 = L50 * 1.1)
head(ret)
```

# Step 13 — Summary and automatic report

`collect_results()` aggregates all results, `summary_table()` produces the
summary table, and `fishstock_report()` generates an HTML / PDF / DOCX report
containing only the available sections.

```{r collect}
res <- collect_results(
  allometry   = lw,
  ypr         = yr,
  tac         = tac,
  quota       = q,
  closure     = clo,
  size_advice = adv,
  maturity    = list(L50 = L50, L95 = L95),
  meta        = list(species = "Cymbium cymbium", stock = "Cymbium")
)
summary_table(res)
```

```{r report, eval=FALSE}
# Full report (requires pandoc; formats: "html", "pdf", "docx")
fishstock_report(res, file = "rapport_cymbium", format = "html",
                 title = "Cymbium cymbium — stock assessment",
                 author = "stockflow")
```

# Pipeline summary

| Step | `stockflow` function(s) | Official package wrapped |
|-------|-------------------------|---------------------------|
| Exploration | `plot_length_frequency`, `plot_length_weight`, `plot_length_evolution` | ggplot2 |
| LFQ | `prepare_tropfish` | TropFishR |
| Growth | `run_powell`, `run_kscan`, `run_elefan`, `run_elefan_ga/sa`, `plot_response_surface`, `plot_growth_modes`, `plot_vbgf` | TropFishR |
| Mortality | `estimate_M_all`, `run_mortality` | FSA, TropFishR, fishmethods |
| Allometry | `fit_length_weight` | FSA |
| Maturity | `plot_maturity_ogive` | stats (GLM) |
| LBSPR | `lbspr_pars`, `lbspr_sim`, `run_lbspr` | LBSPR |
| LBB | `run_lbb_froese`, `lbb_reference_points` | LBB (Froese) |
| YPR | `ypr_param`, `run_ypr`, `ypr_reference_points` | TropFishR |
| TAC / quotas | `compute_tac`, `compute_prorata`, `allocate_quota` | — |
| Biological rest period | `spawning_season`, `optimal_closure` | TropFishR |
| Technical measures | `recommend_sizes`, `retention_curve` | TropFishR |
| Report | `collect_results`, `summary_table`, `fishstock_report` | rmarkdown |

```{r session}
sessionInfo()
```